---
title: "compare maps"
author: 'Noëlle Schenk'
date: "August 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Michel wrote:
Collinearity between the two maps was tested by comparing identical marker positions from both maps

## Collinearity
based on .bed files

"new" map, constructed with PeexOM genome : /home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/2GBS_LebMAP3_ASMap/output_map/F7KAE_Kns.bed

"old" map, constr. on Peexv113 genome and translated to OM genome: /home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/GBS_LepMAP3/output_map/OM_geneticmap.bed
```{r}
require(ggparallel)
```


```{r}
kmm <- read.table("F7AE_Kmm.bed")
kns <- read.table("/home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/2GBS_LebMAP3_ASMap/output_map/F7KAE_Kns.bed")
#kns <- read.table("/home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/GBS_LepMAP3/output_map/OM_geneticmap.bed")


kmm["fullname"] <- paste(paste(kmm$V1, kmm$V2, sep=":"), kmm$V3, sep="-")
kns["fullname"] <- paste(paste(kns$V1, kns$V2, sep=":"), kns$V3, sep="-")
length(which(kns$fullname %in% kmm$fullname))
a <- kns[which(kns$fullname %in% kmm$fullname), ]
b <- kmm[which(kmm$fullname %in% kns$fullname),]
c <- merge(a, b, by="fullname")
d <- c[c("fullname", "V4.x", "V4.y")]

d1 <- data.frame(do.call('rbind', strsplit(as.character(d$V4.x),':',fixed=TRUE)))
d2 <-  data.frame(do.call('rbind', strsplit(as.character(d$V4.y),':',fixed=TRUE)))
dd <- cbind(d1, d2)
names(dd) <- c("gm1lg", "gm1pos", "gm2lg", "gm2pos")

ggparallel(list("gm1lg", "gm2lg"), dd)
colin_plot <- recordPlot()
```
```{r, eval=F}
# store plot
pdf("kmm_kns_bedfile_parallelplot.pdf")
colin_plot
dev.off()
```


# Rqtl comparison
tr.txt contained a list of old_name=new_name, names.txt contained the first row of .csv file where the marker names are stored. This first line was manually exchanged with new.txt
```{bash}
sed -e 's/^/s%/' -e 's/=/%/' -e 's/$/%g/' tests/tr.txt |
sed -f - tests/names.txt > tests/new.txt
```

```{r}
library(qtl2)
library(ASMap)
```
```{r}
# kmm <- read.cross(file="/home/exserta/Documents/master_project_noelle/data/kmm_geneticmap/F7_V3.map.csv", format="csv")
kmm <- read.cross(file="/home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/kns_to_kmm_geneticmaps/tests/new.csv", format="csv")
kns <- read.cross(file="/home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/2GBS_LebMAP3_ASMap/results/geneticmap2.csv", format="csv")

```


wie kann ich maps vergleichen, die nicht auf gleichem genom konstruiert wurden?
marker names übersetzen


Pax nach PexOM


 [976] "Super.Scaffold_11019_205566"  "Super.Scaffold_11019_205540"  "Super.Scaffold_11019_45894"   "Super.Scaffold_11019_149094"  "Super.Scaffold_11019_149221" 
 [981] "Super.Scaffold_11019_149095"  "Super.Scaffold_11019_149098"  "Super.Scaffold_11019_149099"  "Super.Scaffold_11019_149102"  "Super.Scaffold_11019_149214" 
 [986] "Super.Scaffold_11019_149239"  "Super.Scaffold_11019_45893"   "Super.Scaffold_11019_45845"   "Super.Scaffold_11019_45836"   "Super.Scaffold_13920_2182528"
 [991] "Super.Scaffold_13920_2182530" "Super.Scaffold_12246_1069942" "Super.Scaffold_13903_1160005" "Super.Scaffold_13920_468871"  "Super.Scaffold_13889_2389421"
 [996] "Super.Scaffold_12246_1069855" "Super.Scaffold_12200_602716"  "Super.Scaffold_12200_699307"  "Super.Scaffold_12200_699339"  "Super.Scaffold_12200_906446" 
 [ reached getOption("max.print") -- omitted 23 entries ]
> head(markernames(kmm))
[1] "Peaxi162Scf00207_1492281" "Super-Scaffold_13931"     "Super-Scaffold_13931"     "Super-Scaffold_13931"     "Super-Scaffold_13931"     "Super-Scaffold_13931" 

```{r}
data(mapDH, package = "ASMap")

chrl <- sample(c(TRUE,FALSE), 23, replace = TRUE)
mapDH1 <- subset(mapDH, chr = chrl)
alignCross(mapDH, maps = list(DH = mapDH1), layout = c(3,5), col = 1:7)

```

