---
title: "map_construction_ASMap"
author: 'Noëlle Schenk'
date: "August 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Loading required packages, set wd, set dopdf=TRUE.
```{r ,message=F, echo=F}
require(ASMap)
require(RColorBrewer)
require(dplyr)
require(ggplot2)
require(reshape2)
require(ggparallel)
require(wgaim)
setwd("/home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/2GBS_LebMAP3_ASMap")
dopdf = TRUE # to create pdf document
```

## Remove excessive missing
All markers were removed with less than 100 individuals. (= more than 100 individuals missing).

All individuals were removed with less than 1000 markers. (=more than 1000 markers missing).
```{r}
mrks <- read.table("F7-K_OM/LM3_F7_K_OM.MARKER.genoSUMMARY.csv", header=T, sep=",")
mrks <- melt(mrks)
#check for exessive missings
remove_mrks <- mrks[which(mrks$NA. > 100),] # 0, nothing to remove

indvs <- read.table("F7-K_OM/LM3_F7_K_OM.INDIVIDUAL.genoSUMMARY.csv", header=T, sep=",")
remove_inds <- as.character(indvs[which(indvs$NA. > 1000), "individual"]) # are removed in cleaned marker set
```
Individuals `r print(remove_inds)` are removed. 

No markers are removed. `r print(remove_mrks) #0`

```{bash, eval=F}
cat LM3_F7_K_OM.markers.csv | grep -v 'RIL_35' | grep -v 'RIL_44' | grep -v 'RIL_67' | grep -v 'RIL_88' > LM3_F7_K_OM.markers.clean.csv
```
```{r, eval = F}
mc <- read.csv("F7-K_OM/LM3_F7_K_OM.markers.clean.csv")
mc[] <- lapply(mc, as.character)
pheno <- seq(0,ncol(mc)-1)
mc <- rbind(pheno, mc)
# take out parentals
mc <- mc[which(mc$X != "exserta" & mc$X != "axillaris"), ]
write.csv(mc, "F7-K_OM/LM3_F7_K_OM.markers.clean.csv", na = "-", row.names = F)
```
**manually** insert "Genotype" in the upmost left corner and delete entry "0" in second row.

| Genotype | Super.Sc... | Suber.Sc... | ... |
| -------- | ----------- | ----------- | --- |
|          |      1      |      2      | ... |
| exserta  |      AA     |      AA     | ... |
| axillaris|      TT     |      TT     | ... |
|   ...    |     ...     |      ...    | ... |


# Map Construction
## Read in and format data
In the input file, **axillaris** genotype is encoded by "AX", heterozygous as "HET" and **exserta** as "EX". Missing data is encoded as "-". 

| input file | read.cross | f7_K |
|:----|:----|:----|
| AX | AA | 1 |
| HET | AB |  |
| EX | BB | 2 |
| - | | NA |
```{r}
f7_K <- read.cross(format = "csv", file = "F7-K_OM/LM3_F7_K_OM.markers.clean.csv", F.gen = 7, genotypes = c("AX", "HET", "EX"), na.strings = "-")#, crosstype = "riself")
f7_K <- convert2riself(f7_K)
f7_K <- pullCross(f7_K, type = "co.located")
#set bychr = FALSE to allow complete reconstruction of map 
map1 <- mstmap.cross(f7_K, bychr = F, dist.fun = "kosambi", trace = FALSE, detectBadData = F, p.value = 1e-09, mvest.bc = T, return.imputed = T)
# order markers within linkage groups
map1 <- mstmap.cross(map1, bychr = T, dist.fun = "kosambi", trace = FALSE, detectBadData = F, p.value = 1e-09, mvest.bc = F, return.imputed = T)
summary(map1)
```
```{r}
heatMap(map1, lmax=15)
```
```{r}
profileMark(map1, stat.type = c("seg.dist", "dxo", "erf", "lod"), id = "Genotype", layout = c(1, 4), type = "l")
```
```{r}
mean(countXO(map1))
```
```{r}
# push back in markers
map1 <- pushCross(map1, type="co.located")
# order again
map1 <- mstmap(map1, bychr = T, dist.fun = "kosambi", trace = TRUE, detectBadData = F, p.value = 1e-09, mvest.bc = F, return.imputed = T)
heatMap(map1, lmax=12)
```

```{r}
profileMark(map1, stat.type = c("seg.dist", "dxo", "erf", "lod"), id = "Genotype", layout = c(1, 4), type = "l")
```
```{r}
# impute
map1 <- cross2int(map1, id="Genotype", rem.mark=F) # rem.mark = F to not take out colocated markers from the map

# # geno contains the full genetic map
# map1$geno$L.1$map[1:10]
# # imputed contains unique markers
# map1$imputed.geno$L.1$map[1:4]

# merge smallest four linkage groups into one
map1 <- mergeCross(map1, merge = list("L.mixed" = c("L.1", "L.10", "L.11", "L.12" ,"L.9")))
```
```{r}
write.cross(map1, format="csv", filestem="results/geneticmap2")
write.cross(map1, format="qtab", filestem="results/geneticmap2")
# 'qtab' format produces a lot of files I will not further use. Delete them. The only file saved will be 'geneticmap_location.qtab'
file.remove("results/geneticmap2_phenotypes.qtab"); file.remove("results/geneticmap2_founder.qtab"); file.remove("results/geneticmap2_genotypes.qtab"); file.remove("results/geneticmap2_symbols.qtab")
```
To convert the .qtab file to .bed format and take out linkage group 8, run bash: 
```{bash, eval=F}
cat geneticmap2_location.qtab | grep -v '#' | sed 's/Super.Sc/Super-Sc/g' | awk '$2 != "L.mixed"' | tr '_' '\t' | awk '{print $1 "_" $2 "\t" $3 "\t" $3+1 "\t" $4 ":" $5}' > geneticmap2_L1to7.bed
```
```{r}
plot.map(map1, main = "Genetic map, imputed")
```
