---
title: "stats of genomes"
author: 'Noëlle Schenk'
date: "August 17, 2018"
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---
# get assembly stats
This script calculates some stats of a genome in .fasta format.

```{r setup, include=FALSE, echo=F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo = F}
library("dplyr")
require(RColorBrewer)
# display.brewer.all()
# rand.data <- replicate(8,rnorm(100,100,sd=1.5))
# boxplot(rand.data,col=brewer.pal(8,"Set3"))
brcol <- brewer.pal(7, "Dark2")
```

Calculate individual contig lengths from fasta file
```{bash, eval=FALSE}
# nice sctipt, found at http://www.danielecook.com/generate-fasta-sequence-lengths/ - thank you Daniele Cook!
# ${genome_loc} is path-to .fasta file (incl. name), ${genome_name} is name without path
cat ${genome_loc} | awk '$0 ~ ">" {print c; c=0;printf substr($0,2,100) "\t"; } $0 !~ ">" {c+=length($0);} END { print c; }' > ${genome_name}_sizes.txt
```
```{r}
# store path to sizes_txt
Peexv113 <- "/home/exserta/Documents/master_project_noelle/data/exserta_genome_NGS/P.EXSERTA.contigs.v1.1.3_sizes.txt"
PeexOM <- "/home/exserta/Documents/master_project_noelle/data/optical_mapping_raw/P.exserta.opticalmap.v1.fasta.sizes"
PeexOM_kmm2_guo1 <- "/home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/ALLMAPS/aug17_kmm2_guo1/merged.fasta.sizes"
PeexOM_kmm1_guo2 <- "/home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/ALLMAPS/aug21_guo2_kmm1/merged.fasta.sizes" 
PeexOM_kmm_guo <- "/home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/ALLMAPS/aug9_kmm_guo/merged.fasta.sizes"
Peexv20 <- "/home/exserta/Documents/master_project_noelle/data/P.EXSERTA.v.2.0/P.exserta.v.2.0.fasta.sizes"

# CHOOSE YOUR GENOME
sizes_loc <- Peexv20
assembly_name <- "Peexv20"

# initialize stats  list
assemblystats <- list()
# number of chromosomes of organism
nchr <- 7
assemblystats[["assemblyname"]] <- assembly_name
```

## Calc stats
Load contig lengths file and start calculating stats
```{r}
lensFile <- read.table(sizes_loc, col.names = c("contig", "bplength"))
lensFile[,"Mbplength"] <- lensFile$bplength / 1000000
# order by length
lensFile <- arrange(lensFile, desc(Mbplength))
assemblystats["total_lengthMbp"] <- sum(lensFile$Mbplength)
assemblystats["ctgs_gt_10kb"] <- length(which(lensFile$Mbplength > 0.01))
assemblystats["number_of_contigs"] <- nrow(lensFile)
assemblystats["median_ctg_lengthMb"] <- median(lensFile$Mbplength)
assemblystats["longest_ctg"] <- max(lensFile$Mbplength)
```
### count number of ambiguous bases
```{bash, eval=F}
cat /home/exserta/Documents/master_project_noelle/projects/BioNano-exserta/ALLMAPS/aug21_guo2_kmm1/merged.fasta | awk -F"N" '{print NF-1}' | paste -sd+ | bc
```
Gives number in [bp], divide by 1000000 to get [Mbp].

### calculate N50 and L50
keep adding up contigs (first the longest, then the second longest, ...) until you reach 50\% of the genome length. That means, add up contigs until the sum is near to `r lensFile$tot_size/2#772.4826`
```{r}
# calc cumulative sum
lensFile["cumsumMb"] <- cumsum(lensFile$Mbplength)
# find nearest value to 50% of genome length
assemblystats["l50"] <- which.min(abs(lensFile$cumsumMb - assemblystats$total_lengthMbp / 2))
assemblystats["n50"] <- lensFile$Mbplength[assemblystats[["l50"]]]
```
### calculate CN50 and CL50
```{r}
assemblystats[["cl50"]] <- which.min(abs(lensFile$cumsumMb - assemblystats$total_lengthMbp / nchr / 2))
assemblystats[["cn50"]] <- lensFile$Mbplength[assemblystats[["cl50"]]]
```
## save stats
```{r, eval=F}
assemblystats[["lensFile"]] <- lensFile
saveRDS(assemblystats, file=paste("assembly_stats_", assembly_name, sep=""))
```


## plot cumul lengths
```{r}
plot(lensFile$cumsumMb / assemblystats$total_lengthMbp, lensFile$Mbplength, ylab = "contig length [Mbp]", xlab = "cumulative genome length", main=paste("Contig sizes ", assembly_name, sep=""), col=brcol[2], pch=16)
```

# Compare different assemblies
plot contig sizes
```{r}
Peexv113 <- readRDS(file = "assembly_stats_Peexv113")
PeexOM <- readRDS(file="assembly_stats_PeexOM")
PeexOM_kmm2_guo1 <- readRDS(file= "assembly_stats_PeexOM_kmm2_guo1")
PeexOM_kmm1_guo2 <- readRDS(file = "assembly_stats_PeexOM_kmm1_guo2")
PeexOM_kmm_guo <- readRDS(file = "assembly_stats_PeexOM_kmm_guo")
Peexv20 <- readRDS(file= "assembly_stats_Peexv20")
```

```{r}
plot(Peexv113$lensFile$cumsumMb / Peexv113$total_lengthMbp, Peexv113$lensFile$Mbplength, ylab = "contig length [Mbp]", xlab = "cumulative genome length", main="Contig sizes", ylim=c(0,250), col = brcol[1], pch=16)
points(PeexOM$lensFile$cumsumMb / PeexOM$total_lengthMbp, PeexOM$lensFile$Mbplength, col=brcol[2], pch=16)
points(PeexOM_kmm2_guo1$lensFile$cumsumMb / PeexOM_kmm2_guo1$total_lengthMbp, PeexOM_kmm2_guo1$lensFile$Mbplength, col=brcol[3], pch=16)
points(PeexOM_kmm1_guo2$lensFile$cumsumMb / PeexOM_kmm1_guo2$total_lengthMbp, PeexOM_kmm1_guo2$lensFile$Mbplength, col=brcol[4], pch=16)
points(PeexOM_kmm_guo$lensFile$cumsumMb / PeexOM_kmm_guo$total_lengthMbp, PeexOM_kmm_guo$lensFile$Mbplength, col=brcol[5], pch=16)
points(Peexv20$lensFile$cumsumMb / Peexv20$total_lengthMbp, Peexv20$lensFile$Mbplength, col=brcol[6], pch=16)
legend(0.7, 250, legend=c("Peex v.1.1.3", "PeexOM", "PeexOM kmm2 guo1", "PeexOM kmm1 guo2", "PeexOM kmm guo", "Peex v.2.0"), col=brcol[1:6], pch=16, cex=0.8)
```

```{r}
# plot all assemblies
pdf("scaffold_sizes_all.pdf", paper="a4r")
plot(Peexv113$lensFile$cumsumMb / Peexv113$total_lengthMbp, Peexv113$lensFile$Mbplength, ylab = "contig length [Mbp]", xlab = "cumulative genome length", main="Contig sizes", ylim=c(0,250), col = brcol[1], pch=16)
points(PeexOM$lensFile$cumsumMb / PeexOM$total_lengthMbp, PeexOM$lensFile$Mbplength, col=brcol[2], pch=16)
points(PeexOM_kmm2_guo1$lensFile$cumsumMb / PeexOM_kmm2_guo1$total_lengthMbp, PeexOM_kmm2_guo1$lensFile$Mbplength, col=brcol[3], pch=16)
points(PeexOM_kmm1_guo2$lensFile$cumsumMb / PeexOM_kmm1_guo2$total_lengthMbp, PeexOM_kmm1_guo2$lensFile$Mbplength, col=brcol[4], pch=16)
points(PeexOM_kmm_guo$lensFile$cumsumMb / PeexOM_kmm_guo$total_lengthMbp, PeexOM_kmm_guo$lensFile$Mbplength, col=brcol[5], pch=16)
points(Peexv20$lensFile$cumsumMb / Peexv20$total_lengthMbp, Peexv20$lensFile$Mbplength, col=brcol[6], pch=16)
legend(0.7, 250, legend=c("Peex v.1.1.3", "PeexOM", "PeexOM kmm2 guo1", "PeexOM kmm1 guo2", "PeexOM kmm guo", "Peex v.2.0"), col=brcol[1:6], pch=16, cex=0.8)

plot(Peexv113$lensFile$cumsumMb / Peexv113$total_lengthMbp, Peexv113$lensFile$Mbplength, ylab = "contig length [Mbp]", xlab = "cumulative genome length", main="Contig sizes", ylim=c(0,20), col = brcol[1], pch=16)
points(PeexOM$lensFile$cumsumMb / PeexOM$total_lengthMbp, PeexOM$lensFile$Mbplength, col=brcol[2], pch=16)
legend(0.7, 250, legend=c("Peex v.1.1.3", "PeexOM"), col=brcol[1:2], cex=0.8, pch=16)

plot(1, ylab = "contig length [Mbp]", xlab = "cumulative genome length", main="Contig sizes", ylim=c(130,225), xlim=c(0,1), col = "white")
points(PeexOM_kmm2_guo1$lensFile$cumsumMb / PeexOM_kmm2_guo1$total_lengthMbp, PeexOM_kmm2_guo1$lensFile$Mbplength, col=brcol[3], pch=16)
points(PeexOM_kmm1_guo2$lensFile$cumsumMb / PeexOM_kmm1_guo2$total_lengthMbp, PeexOM_kmm1_guo2$lensFile$Mbplength, col=brcol[4], pch=16)
points(PeexOM_kmm_guo$lensFile$cumsumMb / PeexOM_kmm_guo$total_lengthMbp, PeexOM_kmm_guo$lensFile$Mbplength, col=brcol[5], pch=16)
points(Peexv20$lensFile$cumsumMb / Peexv20$total_lengthMbp, Peexv20$lensFile$Mbplength, col=brcol[6], pch=16)
legend(0.7, 230, legend=c("PeexOM kmm2 guo1", "PeexOM kmm1 guo2", "PeexOM kmm guo", "Peex v.2.0"), col=brcol[3:6], cex=0.8, pch=16)
dev.off()
```
```{r}
# plot "old" assemblies only
pdf("scaffold_sizes_old.pdf", paper="a4r")
plot(Peexv113$lensFile$cumsumMb / Peexv113$total_lengthMbp, Peexv113$lensFile$Mbplength, ylab = "contig length [Mbp]", xlab = "cumulative genome length", main="Contig sizes", ylim=c(0,20), col = brcol[1], pch=16)
points(PeexOM$lensFile$cumsumMb / PeexOM$total_lengthMbp, PeexOM$lensFile$Mbplength, col=brcol[2], pch=16)
legend(0.7, 250, legend=c("Peex v.1.1.3", "PeexOM"), col=brcol[1:2], cex=0.8, pch=16)
dev.off()
```
```{r}
# plot "new assemblies only"
pdf("scaffold_sizes_new.pdf", paper="a4r")
plot(1, ylab = "contig length [Mbp]", xlab = "cumulative genome length", main="Contig sizes", ylim=c(130,225), xlim=c(0,1), col = "white")
points(PeexOM_kmm2_guo1$lensFile$cumsumMb / PeexOM_kmm2_guo1$total_lengthMbp, PeexOM_kmm2_guo1$lensFile$Mbplength, col=brcol[3], pch=16)
points(PeexOM_kmm1_guo2$lensFile$cumsumMb / PeexOM_kmm1_guo2$total_lengthMbp, PeexOM_kmm1_guo2$lensFile$Mbplength, col=brcol[4], pch=16)
points(PeexOM_kmm_guo$lensFile$cumsumMb / PeexOM_kmm_guo$total_lengthMbp, PeexOM_kmm_guo$lensFile$Mbplength, col=brcol[5], pch=16)
points(Peexv20$lensFile$cumsumMb / Peexv20$total_lengthMbp, Peexv20$lensFile$Mbplength, col=brcol[6], pch=16)
legend(0.7, 230, legend=c("PeexOM kmm2 guo1", "PeexOM kmm1 guo2", "PeexOM kmm guo", "Peex v.2.0"), col=brcol[3:6], cex=0.8, pch=16)
dev.off()
```
